---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
library(dplyr)
library(lattice)
setwd("E:/Dropbox/Dropbox/Data Science Courses/Data Science Specialization/05 - Reproducible Research/Assignment 1/repdata%2Fdata%2Factivity")
Data = read.csv("activity.csv")
```

The *Data* data frame has the original data loaded and the *Data_DateTransform* data frame has the transformed data with Dates in R format.
```{r}
Data_DateTransform <- Data %>%
  mutate(date=as.Date(date, "%Y-%m-%d"))
```

## What is mean total number of steps taken per day?

```{r}
Data_CC <- Data_DateTransform[complete.cases(Data_DateTransform),]

Data_sum_by_day <- Data_CC %>%
  group_by(date) %>%
  summarise(total=sum(steps))
```

Histogram of the total number of steps taken each day
```{r}
hist(Data_sum_by_day$total)

```

Mean and median number of steps taken each day
```{r}
mean_steps <- mean(Data_sum_by_day$total)
print(mean_steps)

median_steps <- median(Data_sum_by_day$total)
print(median_steps)
```

Mean Number of Steps Per Day: `r as.integer(mean_steps)`
Median Number of Steps Per Day: `r median_steps`

## What is the average daily activity pattern?

Time series plot of the average number of steps taken
```{r}
Data_Interval <- Data_CC %>%
  group_by(interval) %>%
  summarise(average=mean(steps))

plot(x=Data_Interval$interval, y=Data_Interval$average, type='l')

```

The 5-minute interval that, on average, contains the maximum number of steps
```{r}
interval_max = Data_Interval$interval[which.max(Data_Interval$average)]
print(interval_max)
```
The most activity happens around 8:35am.

## Imputing missing values

```{r}
Data_NA <- Data_DateTransform[!complete.cases(Data_DateTransform),]
print(nrow(Data_NA))
missing_values<-nrow(Data_NA)
```
Total Number of Missing Values: `r missing_values`

Fill in the missing values with the mean.
```{r}
df_incomplete <- Data_DateTransform[!complete.cases(Data_DateTransform),]
df_merge <- merge(df_incomplete, Data_Interval)
Data_Impute <- Data_DateTransform
Data_Impute[!complete.cases(Data_Impute),c('steps')] <-df_merge[,c('average')]
```

Histogram of the total number of steps taken each day after missing values are imputed
```{r}
Data_Impute_sum_by_day <- Data_Impute %>%
  group_by(date) %>%
  summarise(total=sum(steps))

hist(Data_Impute_sum_by_day$total)
```


## Are there differences in activity patterns between weekdays and weekends?

```{r}
fun_day_type <- function(x) {
  x <-  as.Date(x)
  day_type <- 'weekday'
  if ( (weekdays(x) == 'Saturday') | (weekdays(x)=='Sunday') ) day_type='weekend'
  day_type
}

Data_Impute$daytype <- apply(Data_Impute, 1, function(x) { fun_day_type(x[2]) } )

df_comp <- Data_Impute %>% 
  group_by(daytype,interval) %>%
  summarise(average=mean(steps))
  
xyplot(average~interval|daytype, data=df_comp, type='l',layout=(c(1,2)))
```

Yes, there is a difference in activity patterns between weekdays and weekends.
